Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane

### 2019-04-02 Manuscript/manuscript.tex
Removing TODOs -- updated figures PC-PS interaction and also with OP-response to ions.

\begin{figure*}[!tbp] 
  \centering 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_kcl.pdf} 
  \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix} 
    CAPTION FOR THE FIGURE WITH POPC: The POPC head group order parameters and the P--N vector angle
    with respect to the membrane normal as a function of POPS content in a bilayer
    from ECC-lipid and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
    Experimental order parameter values are from Ref. \citenum{scherer87}
    and the signs from Ref. \citenum{ferreira16} (only \ce{Na^+} counterions, but shown also in bottom plots).
    CAPTION FOR THE FIGURE WITH POPS: The POPS head group order parameters, the P--N and C$_\beta$--C$_\gamma$ vector angles (stars)
    with respect to the membrane normal as a function of POPC content in a bilayer
    from ECC-lipid and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
    Experimental order parameter values and signs for pure POPS are from Ref. \citenum{NMRlipidsIV} and values
    for POPC:POPS (5:1) mixture from Ref. \citenum{roux90} (only \ce{Na^+} counterions, but shown also in bottom plots).
    %The underestimated response of ECC-lipids with \ce{Na^+} counterions 
    %is likely due to a still slightly overestimated binding affinity of \ce{Na^+} to the phospholipids,
    %which is corroborated by the series with \ce{K^+} counterions (lower affinity than \ce{Na^+}),
    %where the response is closer to the experiment, which uses only \ce{Na^+} counterions. 
  }
  \todo{JOE+SAMULI: split this figure into separate plots for PC and PS, 
    and change the PS-plot x-axis to reflect the dilution of POPS by PC (now shows the content of PS). 
    The experimental lines in the bottom plot shall be different to make clear that they are not with K counterions (gray dashed?).} \\
  \todo{Error bar of 0.02 should be shown for the experimental POPS results because they come from different experiments, see caption of figure 8 in NMRlipids IV publication.}
\end{figure*} 
First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
According to the electrometer concept, the head group order parameters of PC lipids increase
with the increasing amount of negatively charged PS lipids in experiments, because the headgroup P--N dipole
tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
The trend is observed with both force fields in simulations with \ce{K+} counterions and
in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions,
where the order parameters are unchanged or decrease and the P--N vector tilts more perpendicular to the membrane interface.
The result can be explained by the higher binding affinity of sodium in the
Lipid17/Dang ion simulations (Fig~\ref{fig:POPS-counterions-dens}), which
compensates the negative charge of PS lipids at the interface.

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
  \caption{\label{fig:delta_ordPar_NaCl_PCPS} 
    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors
    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
    as a function of \ce{NaCl} concentration from ECC-lipid and Lipid17/Dang simulations at 298 K.
    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
    with the same value for both order parameters such that the lower order
    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
    are shown as lower and upper bounds for the response to \ce{NaCl}, respectively.
    The magnitudes of experimental order parameters are from Ref. \citenum{roux90} and signs from Ref. \citenum{ferreira16}.
    %which has a lower affinity to phospholipid bilayers compared to \ce{LiCl} \citep{roux90}. 
    %The orientation of the \ce{COO^-} group is defined as 
    %the connector from the $\beta$ carbon to the carbon in \ce{COO^-} (stars, bottom right). 
  }
  \todo{We should plot also the experimental results for POPC with KCl in this figure for the lower bound.} \\
  \todo{Why there are no error bars for ECC-POPS in this and other figures?} \\
  \todo{SAMULI: Use the same y-axis scale in the bottom figures showing angles for POPC and POPS to ease the comparison.}
  \end{figure}

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
  \caption{\label{fig:delta_ordPar_KCl_PCPS}
    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors 
    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
    as a function of \ce{KCl} concentration from ECC-lipid and Lipid17/Dang simulations 
    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
    with the same value for both order parameters such that the lower order
    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
  }
  \todo{SAMULI: Use the same y-axis scale in the bottom figures showing angles for POPC and POPS to ease the comparison.}
\end{figure}

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
  \caption{\label{fig:delta_ordPar_CaCl_PCPS} 
    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors 
    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
    as a function of \ce{CaCl} concentration from ECC-lipid and Lipid17/Dang simulations 
    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
    with the same value for both order parameters such that the lower order
    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
  }
  \todo{SAMULI: Use the same y-axis scale in the bottom figures showing angles for POPC and POPS to ease the comparison.}
\end{figure}

>>>>

\begin{figure*}[!tbp] 
  \centering 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_kcl.pdf} 
  \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix} 
    CAPTION FOR THE FIGURE WITH POPC: The POPC head group order parameters and the P--N vector angle
    with respect to the membrane normal as a function of POPS content in a bilayer
    from ECC-lipid and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
    Experimental order parameter values are from Ref. \citenum{scherer87}
    and the signs from Ref. \citenum{ferreira16} (only \ce{Na^+} counterions, but shown also in bottom plots).
    CAPTION FOR THE FIGURE WITH POPS: The POPS head group order parameters, the P--N and C$_\beta$--C$_\gamma$ vector angles (stars)
    with respect to the membrane normal as a function of POPC content in a bilayer
    from ECC-lipid and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
    Experimental order parameter values and signs for pure POPS are from Ref. \citenum{NMRlipidsIV} and values
    for POPC:POPS (5:1) mixture from Ref. \citenum{roux90} (only \ce{Na^+} counterions, but shown also in bottom plots).
    %The underestimated response of ECC-lipids with \ce{Na^+} counterions 
    %is likely due to a still slightly overestimated binding affinity of \ce{Na^+} to the phospholipids,
    %which is corroborated by the series with \ce{K^+} counterions (lower affinity than \ce{Na^+}),
    %where the response is closer to the experiment, which uses only \ce{Na^+} counterions. 
  }
  \todo{JOE+SAMULI: split this figure into separate plots for PC and PS}
\end{figure*} 
First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
According to the electrometer concept, the head group order parameters of PC lipids increase
with the increasing amount of negatively charged PS lipids in experiments, because the headgroup P--N dipole
tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
The trend is observed with both force fields in simulations with \ce{K+} counterions and
in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions,
where the order parameters are unchanged or decrease and the P--N vector tilts more perpendicular to the membrane interface.
The result can be explained by the higher binding affinity of sodium in the
Lipid17/Dang ion simulations (Fig~\ref{fig:POPS-counterions-dens}), which
compensates the negative charge of PS lipids at the interface.

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
  \caption{\label{fig:delta_ordPar_NaCl_PCPS} 
    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors
    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
    as a function of \ce{NaCl} concentration from ECC-lipid and Lipid17/Dang simulations at 298 K.
    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
    with the same value for both order parameters such that the lower order
    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
    are shown as lower and upper bounds for the response to \ce{NaCl}, respectively.
    The magnitudes of experimental order parameters are from Ref. \citenum{roux90} and signs from Ref. \citenum{ferreira16}.
    %which has a lower affinity to phospholipid bilayers compared to \ce{LiCl} \citep{roux90}. 
    %The orientation of the \ce{COO^-} group is defined as 
    %the connector from the $\beta$ carbon to the carbon in \ce{COO^-} (stars, bottom right). 
  }
  \todo{We should plot also the experimental results for POPC with KCl in this figure for the lower bound.} \\
  \end{figure}

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
  \caption{\label{fig:delta_ordPar_KCl_PCPS}
    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors 
    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
    as a function of \ce{KCl} concentration from ECC-lipid and Lipid17/Dang simulations 
    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
    with the same value for both order parameters such that the lower order
    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
  }
\end{figure}

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
  \caption{\label{fig:delta_ordPar_CaCl_PCPS} 
    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors 
    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
    as a function of \ce{CaCl} concentration from ECC-lipid and Lipid17/Dang simulations 
    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
    with the same value for both order parameters such that the lower order
    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
  }
\end{figure}
